### 2023-03-28 draft/draft-daniel-sanchez.Rnw
final commit before turning in draft

While less present in the developed world, informality is very prevalent in developing countries, where it represents a third of low and middle-income countries' economic activity \parencite{Delechat2020}. The work of \textcite{Soto.2002} has motivated much research about informality in Latin America, being one of the regions with the most prevalent levels of informality, as it often is the only option for several workers on the lower end of the wealth distribution \parencite{Oviedo.2009}. At the macroeconomic level, informality could be regarded as one consequence of poor institutions, which have been shown to be significant factors for growth \parencite{Acemoglu.2001, RafaelLaPorta.1997, Glaeser.2004}. Further, its effect on inequality and its potential to be a channel for the propagation of monetary and fiscal policy \textcite{Alberola.2020} make informality a significant macroeconomic determinant of underdevelopment.

The desire for the formalization of the economy is well understood, especially considering that the COVID-19 pandemic has increased informality worldwide \parencite{ILO.2022}. However, the policy angle is subtle, given that the causes and consequences of informality are difficult to document and understand. Some approaches have focused on the role of governments in providing incentives for formalization. Designing tax systems that minimize distortions in the market \parencite{Bardey.2019} as well as reducing the costs of formalization by reducing entry regulation for firms are some of the implementations of this approach. \textcite{MauricioPrado.2011} proposes that firms which are less productive endogenously choose to operate informally due to entry costs and taxation. Thus, reducing entry costs can directly affect formalization of businesses, and an underlying consequence is that these businesses are more likely to hire formal workers.  Figure \ref{fig:fig} displays all active businesses before and after the policy change as well as the amount of job contracts registered under Social Security in Ecuador, showing how employment and business creation tend to move together.

@@

I base my empirical analysis under a theoretical framework informed by the work of \textcite{Branstetter.2014}. In a simple model of entrepreneurship, it is predicted that reduced entry costs will increase firm entry and employment, however, all impacts are \textit{at the margin}: new firms will be generally smaller, less productive and less likely to hire large amounts of new workers. While the results from the empirical analysis show a positive effect of the reform on employment, the effect is not robust to different specifications. Further, event study specifications suggest that due to the timing of the policy, the immediate effects of the policy might be due to catch-up effects from the pandemic rather than due to the underlying economic mechanism of the reform.These results suggest that the true causal effect of the policy on employment is too small to identify with my empirical strategy or even null. Some evidence which supports the validity of the model with Ecuadorian, as the empirical validations in Portugal \parencite{Branstetter.2014} and Mexico \parencite{Kaplan.2011} also find small positive effects of entry cost deregulation. However, the possibility of a null effect of the reform in Ecuador might be explained by the lack of coordination between entry and labour market regulation.

@@

Research on Ecuador is limited and has focused on the effect of exogenous shocks on measurements of informality, notably measuring the effects of a significant earthquake in 2016 \parencite{Mendoza.2020} and the labour market impact of Venezuelan migration \parencite{Olivieri.2022}, both finding positive effects on informality. It is found that poverty and informality are jointly determined, where informal workers often come from low-education and income backgrounds. Further, the informal labour market is mainly supply-led but it is so heterogenously: low-skilled workers often have no choice to work informally, while better educated workers sometimes choose to operate informally \parencite{Canelas.2019}. On the firm side, it is determined that the endogenous choice of sector is also present in Ecuador \textcite{Medvedev.2016}. Ecuador also documents a formal sector wage premium \textcite{MacIsaac.1997}, which, according to \textcite{ElianeElBadaoui.2010}, responds to a firm size wage premium due to the endogenous choice of the firm to operate in the informal sector.

@@

Future research should perhaps focus on controlling for omitted variables as much as possibile, using additional data available at a monthly frequency and at the province level. Some possibilities are to include Google mobility reports available since 2020, however, it would restrict the sample to pandemic years only, which may render conclusions which are not externally valid. Further, mobility data may overcontrol for factors and thus not work for the identification of the policy, however, these data should be useful to perform robustness checks for the soundness of the model. Additionally, a better approach to the study of this event study without a control group is time series forecasting, given that the empirical analysis found evidence of an autoregresive process. Using ARIMA or SARIMAX models may prove useful for the identification of the causal effect, however, once again the reduced sample problem may happen.

>>>>

While less present in the developed world, informality is very prevalent in developing countries, where it represents a third of low and middle-income countries' economic activity \parencite{Delechat2020}. The work of \textcite{Soto.2002} has motivated much research about informality in Latin America, being one of the regions with the most prevalent levels of informality, as it often is the only option for several workers on the lower end of the wealth distribution \parencite{Oviedo.2009}. At the macroeconomic level, informality could be regarded as one consequence of poor institutions, which have been shown to be significant factors for growth \parencite{Acemoglu.2001, RafaelLaPorta.1997, Glaeser.2004}. Further, its effect on inequality and its potential to be a channel for the propagation of monetary and fiscal policy \parencite{Alberola.2020} make informality a significant macroeconomic determinant of underdevelopment.

The desire for the formalization of the economy is well understood, especially considering that the COVID-19 pandemic has increased informality worldwide \parencite{ILO.2022}. However, the policy angle is subtle, given that the causes and consequences of informality are difficult to document and understand. Some approaches have focused on the role of governments in providing incentives for formalization. Designing tax systems that minimize distortions in the market \parencite{Bardey.2019} as well as reducing the costs of formalization by reducing entry regulation for firms are some of the implementations of this approach. \textcite{MauricioPrado.2011} proposes that firms which are less productive endogenously choose to operate informally due to entry costs and taxation. Thus, reducing entry costs can directly affect formalization of businesses and an underlying consequence is that these businesses are more likely to hire formal workers. Figure \ref{fig:fig} displays all active businesses before and after the policy change as well as the amount of job contracts registered under Social Security in Ecuador, showing how employment and business creation tend to move together.

@@

I base my empirical analysis under a theoretical framework informed by the work of \textcite{Branstetter.2014}. In a simple model of entrepreneurship, it is predicted that reduced entry costs will increase firm entry and employment, however, all impacts are \textit{at the margin}: new firms will be generally smaller, less productive and less likely to hire large amounts of new workers. While the results from the empirical analysis show a positive effect of the reform on employment, the effect is not robust to different specifications. Further, event study specifications suggest that due to the timing of the policy, the immediate effects of the policy might be due to catch-up effects from the pandemic rather than due to the underlying economic mechanism of the reform.These results suggest that the true causal effect of the policy on employment is too small to identify with my empirical strategy or even null. Some evidence which supports the validity of the model with Ecuadorian data, as the empirical validations in Portugal \parencite{Branstetter.2014} and Mexico \parencite{Kaplan.2011} also find small positive effects of entry cost deregulation. However, the possibility of a null effect of the reform in Ecuador might be explained by the lack of coordination between entry and labour market regulation.

@@

Research on Ecuador is limited and has focused on the effect of exogenous shocks on measurements of informality, notably measuring the effects of a significant earthquake in 2016 \parencite{Mendoza.2020} and the labour market impact of Venezuelan migration \parencite{Olivieri.2022}, both finding positive effects on informality. It is found that poverty and informality are jointly determined, where informal workers often come from low education and low income backgrounds. Further, the informal labour market is mainly supply-led but it is so heterogenously: low-skilled workers often have no choice to work informally, while better educated workers sometimes choose to operate informally \parencite{Canelas.2019}. On the firm side, it is determined that the endogenous choice of sector is also present in Ecuador \parencite{Medvedev.2016}. Ecuador also documents a formal sector wage premium \textcite{MacIsaac.1997}, which, according to \textcite{ElianeElBadaoui.2010}, responds to a firm size wage premium due to the endogenous choice of the firm to operate in the informal sector.

@@

Future research should perhaps focus on controlling for omitted variables as much as possible, using additional data available at a monthly frequency and at the province level. Some possibilities are to include Google mobility reports available since 2020, however, it would restrict the sample to pandemic years only, which may render conclusions which are not externally valid. Further, mobility data may overcontrol for factors and thus not work for the identification of the policy, however, these data should be useful to perform robustness checks for the soundness of the model. Additionally, a better approach to the study of this event study without a control group is time series forecasting, given that the empirical analysis found evidence of an autoregresive process. Using ARIMA or SARIMAX models may prove useful for the identification of the causal effect, however, once again the reduced sample problem may happen.
